---
title: "Drug Consumption Data Analysis"
output: github_document
---

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(viridis)
library(hrbrthemes)
library(RColorBrewer)
library(ggjoy)
library(cowplot)
library(corrplot)
library(rcompanion)
library(corrr)

```

> ## load dataset

```{r}
data = read.csv("drug.csv")
head(data, n=3)
```

> ## give factor levels to education

```{r}
data$Education = factor(data$Education, levels=c("Doctorate Degree", "Masters Degree",
                                "University Degree", "Professional Certificate/ Diploma",
                                "Some College,No Certificate Or Degree", "Left School at 18 years",
                                "Left School at 17 years", "Left School at 16 years",
                                "Left School Before 16 years"))
```

> ## recode usage levels in drugs

```{r}
data = data %>%
  mutate(across(c(14:32), ~ recode(., 
                                'CL0'= 0,
                                'CL1'= 1,
                                'CL2'= 2,
                                'CL3'= 3,
                                'CL4'= 4,
                                'CL5'= 5,
                                'CL6'= 6,
                )))

```

> ## group column names

```{r}
drug.cols = colnames(data[, 14:32])
score.cols = colnames(data[, 7:13])
feature.cols = colnames(data[, 1:6])
```

> ## create drug types

```{r}
stimulants <- c("Amphet", "Caff", "Coke", "Crack", "Meth", "Nicotine")
depressants <- c("Alcohol", "Benzos", "Heroin", "Legalh", "Amyl", "VSA")
hallucinogens <- c("LSD", "Mushrooms", "Ketamine", "Cannabis", "Ecstasy", "Choc")

```

> ## create seperate datasets for features and labels

```{r}
labels = data[, 14:32]
features = data[, 1:13]
```

> ## check the usage level (addiction) of each drug

```{r}
drug.tb = gather(labels, key="drug")

ggplot(drug.tb, aes(x=value)) +
  geom_bar(aes(fill=as.factor(value))) +
  facet_wrap(~drug) +
  scale_fill_brewer(palette="Spectral") +
  theme(legend.position="bottom")
  
```

> ## create a dataframe **temp** of indivuals who actively uses dangerous drugs

```{r warning=FALSE}
drugs = colnames(labels)
dng.drugs = c("Heroin", "Meth", "Crack", "Coke")
recent.usage = c(4, 5, 6)

temp = data %>%
  filter(Heroin %in% recent.usage | Meth %in% recent.usage | Crack %in% recent.usage | Coke %in% recent.usage)

```

> ## plot the distribution in education levels in **temp**

```{r}
ggplot(temp, aes(y=Education)) +
  geom_bar(aes(fill=Education)) +
  scale_fill_brewer(palette="RdBu") +
  labs(title="dangerous drug usage in education groups")

```

> ## plot the distribution in age levels in **temp**

```{r}
ggplot(temp, aes(y=Age)) +
  geom_bar(aes(fill=Age)) + 
  scale_fill_ipsum() +
  labs(title="dagerous drugs usage in age groups")

```

> ## Nscore behaviour in drug groups

```{r}
df.score = labels %>%
  bind_cols(data[, 7:13])

df.score.pivot = gather(df.score, key="drug", "usage", 1:19)

df.score.pivot %>%
  filter(usage >= 4) %>%
  filter(drug %in% dng.drugs) %>%
  ggplot(., aes(x=Nscore)) + 
    geom_histogram(bins=20) +
    facet_wrap(~drug)

```

> ## score distributions in depressant users

```{r}

color = brewer.pal(5, "Spectral")

for (i in hallucinogens) {
  p = gather(data, key="score", value="value", score.cols) %>%
    filter(!!sym(i) >=4) %>%
    ggplot(., aes(x=value)) +
    geom_histogram(bins=20, aes(y=..density.., fill="count")) +
    geom_density(aes(fill="density", alpha=0.5)) +
    labs(title=i) +
    facet_wrap(~score) +
    scale_fill_brewer(palette="Set1")
  print(p)
}


```

> ## Analysis on Cannabis
> #### age group distribution in each usage group

```{r}
i = c("Cannabis")
data %>%
  group_by(!!sym(i), Age) %>%
  tally() %>%
  ggplot(., aes(y=Age, x=n, fill=Age)) +
  geom_bar(stat="identity") +
  scale_fill_ft() +
  facet_wrap(vars(!!sym(i)))
  
  
```


> ## popular Ages groups in each drug

```{r}
plt1 = data %>%
  gather(., key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  mutate(drug = fct_reorder(.f = drug,
                            .x = Age,
                            .fun = function(.x) mean(.x == "18 - 24"),
                            .desc = T)) %>%
  ggplot(., aes(y=drug, fill=Age)) +
  geom_bar(position="fill") +
  scale_fill_brewer(palette="Accent") +
  theme(legend.position="bottom")

age.popular.drugs = data %>%
  gather(., key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  group_by(Age, drug) %>%
  tally() %>%
  group_by(Age) %>%
  slice_max(order_by=n, n=7) %>%
  ungroup() %>%
  distinct(drug)

plt2 = data %>%
  gather(., key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  mutate(drug = if_else(drug %in% c(age.popular.drugs)$drug, drug, "other")) %>%
  ggplot(., aes(y=Age, fill=drug)) +
  geom_bar(position="fill") +
  scale_fill_brewer(palette="Paired") +
  theme(legend.position="bottom")

plot_grid(plt1, plt2, ncol=2, rel_heights=c(2, 1))
print(plt1)
print(plt2)

```
> ## popular drugs in education levels

```{r}


education.popular.drugs = data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  group_by(Education, drug) %>%
  tally() %>%
  slice_max(order_by=n, n=7) %>%
  ungroup() %>%
  distinct(drug) %>%
  pull()

data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  mutate(drug = if_else(drug %in% education.popular.drugs, drug, "other")) %>%
  ggplot(aes(y=Education, fill=drug)) +
  geom_bar(position="fill") +
  scale_fill_brewer(palette="Paired")
  

```

> ## popular drugs in Countries

```{r}

country.popular.drugs = data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  group_by(Country, drug) %>%
  tally() %>%
  slice_max(order_by=n, n=7) %>%
  ungroup() %>%
  distinct(drug) %>%
  pull()

data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  mutate(drug = if_else(drug %in% country.popular.drugs, drug, "other")) %>%
  ggplot(aes(y=Country, fill=drug)) +
  geom_bar(position="fill") +
  scale_fill_brewer(palette="Paired")

```

> legal constraint in countries

```{r}
# Group 1: Low Perceived Risk
low_perceived_risk <- c("Alcohol", "Caff", "Choc")

# Group 2: Moderate Perceived Risk
moderate_perceived_risk <- c("Nicotine", "Benzos", "Amphet", "Ketamine", "Meth", "Cannabis")

# Group 3: High Perceived Risk
high_perceived_risk <- c("Coke", "Crack", "Ecstasy", "Heroin", "LSD", "Mushrooms", "Amyl", "Legalh", "Semer", "VSA")

data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  mutate(drug.risk = case_when(drug %in% low_perceived_risk ~ "low",
                               drug %in% moderate_perceived_risk ~ "moderate",
                               drug %in% high_perceived_risk ~ "high",
                               T ~ "unknown")) %>%
  mutate(drug.risk = factor(drug.risk, levels=c("low", "moderate", "high"))) %>%
  mutate(Country = fct_reorder(.f = Country,
                               .x = drug.risk,
                               .fun = function(.x) mean(.x == "high"),
                               .desc = T)) %>%
  ggplot(aes(y=Country, fill=drug.risk)) +
  geom_bar(position="fill") +
  scale_fill_brewer(direction = 1)
  


```

> ## popular drugs within genders

```{r}

lvls <- data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  count(drug, Gender) %>%
  group_by(drug) %>%
  mutate(prop = prop.table(n)) %>%
  filter(Gender == "Female") %>%
  arrange(prop) %>%
  pull(drug) %>%
  as.character()

data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  mutate(drug = factor(drug, levels=lvls)) %>%
  ggplot(aes(y=drug, fill=Gender)) +
  geom_bar(position="fill")

data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  mutate(drug=fct_reorder(.f = drug, 
                          .x = Gender,
                          .fun = function(.x) mean(.x == "Female"),
                          .desc = TRUE)) %>%
  ggplot(aes(y=drug, fill=Gender)) +
  geom_bar(position="fill") +
  scale_fill_brewer(palette="Pastel1")

```

> ## corealtion among drugs

```{r}
data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  spread(key="drug", value="usage") %>%
  mutate(across(drug.cols, ~replace_na(., 0))) %>%
  select(all_of(drug.cols)) %>%
  cor(., use = "complete.obs") %>%
  corrplot(., method = "square", type = "lower", order = 'FPC', tl.col = "black", tl.srt = 45, diag = F)

data %>%
  select(all_of(drug.cols)) %>%
  cor(., use = "complete.obs") %>%
  corrplot(., method = "square", type = "lower", order = 'FPC', tl.col = "black", tl.srt = 45, diag = F)

v1 <- data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  mutate(usage = if_else(usage <= 4, "No", "Yes")) %>%
  mutate(usage = factor(usage)) %>%
  spread(key="drug", value="usage") %>%
  select(all_of(drug.cols)) %>%
  select(-all_of("Semer")) %>%
  colpair_map(., cramerV)

v1[, -1][is.na(v1[, -1])] = 1
mat = as.matrix(v1[, -1])
colnames(mat) = colnames(v1[, -1])
rownames(mat) = colnames(v1[, -1])

corrplot(mat, method = "square", type = "lower", order = 'FPC', tl.col = "black", tl.srt = 45, diag = F)


```



```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

```{r}
display.brewer.all()
brewer.pal.info
```
