---
title: "Drug Consumption Data Analysis"
output: github_document
---

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(viridis)
library(hrbrthemes)
library(RColorBrewer)
library(ggjoy)

```

> ## load dataset

```{r}
data = read.csv("drug.csv")
head(data, n=3)
```

> ## give factor levels to education

```{r}
data$Education = factor(data$Education, levels=c("Doctorate Degree", "Masters Degree",
                                "University Degree", "Professional Certificate/ Diploma",
                                "Some College,No Certificate Or Degree", "Left School at 18 years",
                                "Left School at 17 years", "Left School at 16 years",
                                "Left School Before 16 years"))
```

> ## recode usage levels in drugs

```{r}
data = data %>%
  mutate(across(c(14:32), ~ recode(., 
                                'CL0'= 0,
                                'CL1'= 1,
                                'CL2'= 2,
                                'CL3'= 3,
                                'CL4'= 4,
                                'CL5'= 5,
                                'CL6'= 6,
                )))

```

> ## group column names

```{r}
drug.cols = colnames(data[, 14:32])
score.cols = colnames(data[, 7:13])
feature.cols = colnames(data[, 1:6])
```

> ## create drug types

```{r}
stimulants <- c("Amphet", "Caff", "Coke", "Crack", "Meth", "Nicotine")
depressants <- c("Alcohol", "Benzos", "Heroin", "Legalh", "Amyl", "VSA")
hallucinogens <- c("LSD", "Mushrooms", "Ketamine", "Cannabis", "Ecstasy", "Choc")

```

> ## create seperate datasets for features and labels

```{r}
labels = data[, 14:32]
features = data[, 1:13]
```

> ## check the usage level (addiction) of each drug

```{r}
drug.tb = gather(labels, key="drug")

ggplot(drug.tb, aes(y=value)) +
  geom_bar() +
  facet_wrap(~drug) +
  scale_color_ipsum()
  
```

> ## create a dataframe **temp** of indivuals who actively uses dangerous drugs

```{r warning=FALSE}
drugs = colnames(labels)
dng.drugs = c("Heroin", "Meth", "Crack", "Coke")
recent.usage = c(4, 5, 6)

temp = data %>%
  filter(Heroin %in% recent.usage | Meth %in% recent.usage | Crack %in% recent.usage | Coke %in% recent.usage)

```

> ## plot the distribution in education levels in **temp**

```{r}
ggplot(temp, aes(y=Education)) +
  geom_bar(aes(fill=Education)) +
  scale_fill_brewer(palette="RdBu") +
  labs(title="dangerous drug usage in education groups")

```

> ## plot the distribution in age levels in **temp**

```{r}
ggplot(temp, aes(y=Age)) +
  geom_bar(aes(fill=Age)) + 
  scale_fill_ft() +
  labs(title="dagerous drugs usage in age groups")

```

> ## Nscore behaviour in drug groups

```{r}
df.score = labels %>%
  bind_cols(data[, 7:13])

df.score.pivot = gather(df.score, key="drug", "usage", 1:19)

df.score.pivot %>%
  filter(usage >= 4) %>%
  filter(drug %in% dng.drugs) %>%
  ggplot(., aes(x=Nscore)) + 
    geom_histogram(bins=20) +
    facet_wrap(~drug)

```

> ## score distributions in depressant users

```{r}

color = brewer.pal(5, "Spectral")

for (i in hallucinogens) {
  p = gather(data, key="score", value="value", score.cols) %>%
    filter(!!sym(i) >=4) %>%
    ggplot(., aes(x=value)) +
    geom_histogram(bins=20, aes(y=..density.., fill="count")) +
    geom_density(aes(fill="density", alpha=0.5)) +
    labs(title=i) +
    facet_wrap(~score) +
    scale_fill_brewer(palette="Set1")
  print(p)
}


```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}

display.brewer.all()
```
