---
title: "Drug Consumption Data Analysis"
output: github_document
---

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(viridis)
library(hrbrthemes)
library(RColorBrewer)
library(ggjoy)
library(cowplot)
library(corrplot)
library(rcompanion)
library(corrr)
library(FactoMineR)
library(factoextra)
library(caret)

```

> ## load dataset

```{r}
data = read.csv("drug.csv")

head(data, n=3)
```

> ## give factor levels to education

```{r}
data$Education = factor(data$Education, levels=c("Doctorate Degree", "Masters Degree",
                                "University Degree", "Professional Certificate/ Diploma",
                                "Some College,No Certificate Or Degree", "Left School at 18 years",
                                "Left School at 17 years", "Left School at 16 years",
                                "Left School Before 16 years"), ordered=T)
```

> ## recode usage levels in drugs

```{r}
data = data %>%
  mutate(across(c(13:32), ~ recode(.,
                                'CL0'= 0,
                                'CL1'= 1,
                                'CL2'= 2,
                                'CL3'= 3,
                                'CL4'= 4,
                                'CL5'= 5,
                                'CL6'= 6,
                )))

data <- data %>%
  mutate(Age = case_when(
    Age %in% c("55 - 64", "65+") ~ "55+",
    TRUE ~ Age
  ))

```


```{r}
data <- data %>%
  filter(Semer == 0)
  
trainIndex <- createDataPartition(data$Alcohol, p=0.8, times=1, list=F)
data <- data[trainIndex, ]
val <- data[-trainIndex, ]

data <- data %>%
  select(-all_of(c("Semer", "X")))
  
```


```{r}

```




> ## group column names

```{r}
drug.cols = colnames(data[, 13:30])
score.cols = colnames(data[, 6:12])
feature.cols = colnames(data[, 1:5])

```

> ## create drug types

```{r}
# stimulants <- c("Amphet", "Caff", "Coke", "Crack", "Meth", "Nicotine")
# depressants <- c("Alcohol", "Benzos", "Heroin", "Legalh", "Amyl", "VSA")
# hallucinogens <- c("LSD", "Mushrooms", "Ketamine", "Cannabis", "Ecstasy", "Choc")

stimulants <- c("Amphet", "Coke", "Crack", "Meth", "Nicotine")
depressants <- c("Alcohol", "Benzos", "Heroin", "Legalh", "Amyl", "VSA")
hallucinogens <- c("LSD", "Mushrooms", "Ketamine", "Cannabis", "Ecstasy")

```

> ## create seperate datasets for features and labels

```{r}
labels = data[, 13:30]
features = data[, 1:12]

df <- data %>%
  mutate_at(vars(score.cols), ~ cut(.,
                                    breaks=c(-Inf, -1, 0, 1, Inf),
                                    labels=c("Very Low", "Low", "High", "Very High"),
                                    ))

```

> ## check the usage level (addiction) of each drug

```{r}
drug.tb = gather(labels, key="drug")

ggplot(drug.tb, aes(x=value)) +
  geom_bar(aes(fill=as.factor(value))) +
  facet_wrap(~drug) +
  scale_fill_brewer(palette="Spectral") +
  theme(legend.position="bottom")
  
```

> ## create a dataframe **temp** of indivuals who actively uses dangerous drugs

```{r warning=FALSE}
drugs = colnames(labels)
dng.drugs = c("Heroin", "Meth", "Crack", "Coke")
recent.usage = c(4, 5, 6)

temp = data %>%
  filter(Heroin %in% recent.usage | Meth %in% recent.usage | Crack %in% recent.usage | Coke %in% recent.usage)

```

> ## plot the distribution in education levels in **temp**

```{r}
ggplot(temp, aes(y=Education)) +
  geom_bar(aes(fill=Education)) +
  scale_fill_brewer(palette="RdBu") +
  labs(title="dangerous drug usage in education groups")



```

> ## plot the distribution in age levels in **temp**

```{r}
ggplot(temp, aes(y=Age)) +
  geom_bar(aes(fill=Age)) + 
  scale_fill_ipsum() +
  labs(title="dagerous drugs usage in age groups")

```

> ## Nscore behaviour in drug groups

```{r}
df.score = labels %>%
  bind_cols(data[, score.cols])

df.score.pivot = gather(df.score, key="drug", "usage", drug.cols)

df.score.pivot %>%
  filter(usage >= 4) %>%
  filter(drug %in% dng.drugs) %>%
  ggplot(., aes(x=Nscore)) + 
    geom_histogram(bins=20) +
    facet_wrap(~drug)

```

> ## score distributions in depressant users

```{r}

color = brewer.pal(5, "Spectral")

for (i in stimulants) {
  p = gather(data, key="score", value="value", score.cols) %>%
    filter(!!sym(i) >=4) %>%
    ggplot(., aes(x=value)) +
    geom_histogram(bins=20, aes(y=..density.., fill="count")) +
    geom_density(aes(fill="density", alpha=0.5)) +
    labs(title=i) +
    facet_wrap(~score) +
    scale_fill_brewer(palette="Set1")
  print(p)
}


```

> ## Analysis on Cannabis
> #### age group distribution in each usage group

```{r}
i = c("Heroin")

data %>%
  mutate_at(vars(score.cols), ~ cut(.,
                                    breaks=c(-Inf, -1, 0, 1, Inf),
                                    labels=c("Very Low", "Low", "High", "Very High"),
                                    )
  ) %>%
  gather(key="score", value="level", score.cols) %>%
  mutate(level = factor(level, levels=c("Very Low", "Low", "High", "Very High"))) %>%
  filter(!!sym(i) >= 4) %>%
  ggplot(., aes(y=level, fill=level)) +
  geom_bar() +
  facet_wrap(~score)

data %>%
  mutate(group = ifelse(!!sym(i) >= 4, "active", "inactive"))%>%
  ggplot(., aes(x=Impulsive, y=group, fill=group)) +
    geom_joy(scale=1, alpha=0.7, rel_min_height=0.02)

data %>%
  mutate(group = ifelse(!!sym(i) >= 4, "active", "inactive"))%>%
  ggplot(., aes(x=SS, y=group, fill=group)) +
    geom_joy(scale=1, alpha=0.7, rel_min_height=0.02)

data %>%
  mutate(group = ifelse(!!sym(i) >= 4, "active", "inactive")) %>%
  ggplot(., aes(y=Education, fill=group)) +
  geom_bar(position="fill")

data %>%
  mutate(group = ifelse(!!sym(i) >= 4, "active", "inactive")) %>%
  ggplot(., aes(y=Age, fill=group)) +
  geom_bar(position="fill")

data %>%
  mutate(group = ifelse(!!sym(i) >= 4, "active", "inactive")) %>%
  mutate(Country=fct_reorder(.f = Country, 
                          .x = group,
                          .fun = function(.x) mean(.x == "active"),
                          .desc = TRUE)) %>%
  ggplot(., aes(y=Country, fill=group)) +
  geom_bar(position="fill")
  
```

> ## popular drugs within high Escore individuals

```{r}
i = "Escore"

drug.cols.new = drug.cols[!drug.cols %in% c("Caff", "Choc", "Alcohol", "Nicotine", "Cannabis")]

for (i in score.cols) {
  score.popular.drugs <- df %>%
    gather(., key="drug", value="usage", drug.cols.new) %>%
    filter(usage >= 4) %>%
    filter(!!sym(i) %in% c("High", "Very High")) %>%
    group_by(!!sym(i), drug) %>%
    tally() %>%
    group_by(!!sym(i)) %>%
    slice_max(order_by=n, n=7) %>%
    ungroup() %>%
    distinct(drug)
  
  plt <- df %>%
    gather(., key="drug", value="usage", drug.cols.new) %>%
    filter(usage >= 4) %>%
    mutate(drug = if_else(drug %in% c(score.popular.drugs)$drug, drug, "other")) %>%
    ggplot(., aes(y=!!sym(i), fill=drug)) +
    geom_bar(position="fill") +
    scale_fill_brewer(palette="Paired") +
    theme(legend.position="bottom")
  
  print(plt)
}
  

```


```{r}
```


```{r}
```


```{r}

```

```{r}

```


```{r}
data %>%
  mutate(Education = as.factor(as.numeric(factor(Education)))) %>%
  ggplot(., aes(y=Education, x=Escore, fill=Education)) +
  geom_joy(scale=1, alpha=0.7, rel_min_height=0.02) +
  scale_fill_manual(labels=levels(data$Education), values=brewer.pal(9, "PuRd"))

```


```{r}
# data %>%
#   filter(Crack >= 4) %>%
#   select(all_of(score.cols)) %>%
#   cor(., use = "complete.obs") %>%
#   corrplot(., method = "square", type = "lower", order = 'alphabet', tl.col = "black", tl.srt = 45, diag = F)

data %>%
  select(all_of(score.cols)) %>%
  cor(., use = "complete.obs") %>%
  corrplot(., method = "square", type = "lower", order = 'alphabet', tl.col = "black", tl.srt = 45, diag = F)

```



```{r}
ethnicity.popular.drugs = data %>%
  mutate(Ethnicity = if_else(Ethnicity == "White", "White", "Other")) %>%
  gather(., key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  group_by(Ethnicity, drug) %>%
  tally() %>%
  group_by(Ethnicity) %>%
  slice_max(order_by=n, n=5) %>%
  ungroup() %>%
  distinct(drug)


data %>%
  mutate(Ethnicity = if_else(Ethnicity == "White", "White", "Other")) %>%
  gather(key="drug", value="usage", drug.cols) %>% 
  filter(usage >=4) %>%
  mutate(drug = if_else(drug %in% c(ethnicity.popular.drugs)$drug, drug, "other")) %>%
  ggplot(., aes(x=Ethnicity, fill=drug)) +
  geom_bar(position="fill") +
  scale_fill_brewer(palette="Spectral") +
  theme(legend.position="bottom")

  
```


> ## popular Ages groups in each drug

```{r fig.align="center", echo = FALSE,fig.width = 10}
plt1 = data %>%
  gather(., key="drug", value="usage", drug.cols.new) %>%
  filter(usage >= 4) %>%
  mutate(drug = fct_reorder(.f = drug,
                            .x = Age,
                            .fun = function(.x) mean(.x == "18 - 24"),
                            .desc = T)) %>%
  ggplot(., aes(y=drug, fill=Age)) +
  geom_bar(position="fill") +
  scale_fill_brewer(palette="Accent") +
  theme(legend.position="bottom")

age.popular.drugs = data %>%
  gather(., key="drug", value="usage", drug.cols.new) %>%
  filter(usage >= 4) %>%
  group_by(Age, drug) %>%
  tally() %>%
  group_by(Age) %>%
  slice_max(order_by=n, n=7) %>%
  ungroup() %>%
  distinct(drug)

plt2 = data %>%
  gather(., key="drug", value="usage", drug.cols.new) %>%
  filter(usage >= 4) %>%
  mutate(drug = if_else(drug %in% c(age.popular.drugs)$drug, drug, "other")) %>%
  ggplot(., aes(y=Age, fill=drug)) +
  geom_bar(position="fill") +
  scale_fill_brewer(palette="Paired") +
  theme(legend.position="bottom")

plot_grid(plt1, plt2, ncol=2, rel_heights=c(2, 1))
print(plt1)
print(plt2)

```
> ## popular drugs in education levels

```{r}


education.popular.drugs = data %>%
  gather(key="drug", value="usage", drug.cols.new) %>%
  filter(usage >= 4) %>%
  group_by(Education, drug) %>%
  tally() %>%
  slice_max(order_by=n, n=5) %>%
  ungroup() %>%
  distinct(drug) %>%
  pull()

data %>%
  gather(key="drug", value="usage", drug.cols.new) %>%
  filter(usage >= 4) %>%
  mutate(drug = if_else(drug %in% education.popular.drugs, drug, "other")) %>%
  ggplot(aes(y=Education, fill=drug)) +
  geom_bar(position="fill") +
  scale_fill_brewer(palette="Paired")
  

```

> ## popular drugs in Countries

```{r}

country.popular.drugs = data %>%
  gather(key="drug", value="usage", drug.cols.new) %>%
  filter(usage >= 4) %>%
  group_by(Country, drug) %>%
  tally() %>%
  slice_max(order_by=n, n=6) %>%
  ungroup() %>%
  distinct(drug) %>%
  pull()
country.popular.drugs

data %>%
  gather(key="drug", value="usage", drug.cols.new) %>%
  filter(usage >= 4) %>%
  mutate(drug = if_else(drug %in% country.popular.drugs, drug, "other")) %>%
  ggplot(aes(y=Country, fill=drug)) +
  geom_bar(position="fill") +
  scale_fill_brewer(palette="Paired")

```

> ## legal constraint in countries

```{r}
# Group 1: Low Perceived Risk
low_perceived_risk <- c("Alcohol", "Caff", "Choc")

# Group 2: Moderate Perceived Risk
moderate_perceived_risk <- c("Nicotine", "Benzos", "Amphet", "Ketamine", "Meth", "Cannabis")

# Group 3: High Perceived Risk
high_perceived_risk <- c("Coke", "Crack", "Ecstasy", "Heroin", "LSD", "Mushrooms", "Amyl", "Legalh", "Semer", "VSA")

data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  mutate(drug.risk = case_when(drug %in% low_perceived_risk ~ "low",
                               drug %in% moderate_perceived_risk ~ "moderate",
                               drug %in% high_perceived_risk ~ "high",
                               T ~ "unknown")) %>%
  mutate(drug.risk = factor(drug.risk, levels=c("low", "moderate", "high"))) %>%
  mutate(Country = fct_reorder(.f = Country,
                               .x = drug.risk,
                               .fun = function(.x) mean(.x == "high"),
                               .desc = T)) %>%
  ggplot(aes(y=Country, fill=drug.risk)) +
  geom_bar(position="fill") +
  scale_fill_brewer(direction = 1)
  


```

> ## popular drugs within genders

```{r}

lvls <- data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  count(drug, Gender) %>%
  group_by(drug) %>%
  mutate(prop = prop.table(n)) %>%
  filter(Gender == "Female") %>%
  arrange(prop) %>%
  pull(drug) %>%
  as.character()

data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  mutate(drug = factor(drug, levels=lvls)) %>%
  ggplot(aes(y=drug, fill=Gender)) +
  geom_bar(position="fill")

data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  mutate(drug=fct_reorder(.f = drug, 
                          .x = Gender,
                          .fun = function(.x) mean(.x == "Female"),
                          .desc = TRUE)) %>%
  ggplot(aes(y=drug, fill=Gender)) +
  geom_bar(position="fill") +
  scale_fill_brewer(palette="Pastel1")

```

> ## corealtion among drugs

```{r}
data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  filter(usage >= 4) %>%
  spread(key="drug", value="usage") %>%
  mutate(across(drug.cols, ~replace_na(., 0))) %>%
  select(all_of(drug.cols)) %>%
  cor(., use = "complete.obs") %>%
  corrplot(., method = "square", type = "lower", order = 'FPC', tl.col = "black", tl.srt = 45, diag = F)

data %>%
  select(all_of(drug.cols)) %>%
  cor(., use = "complete.obs") %>%
  corrplot(., method = "square", type = "lower", order = 'FPC', tl.col = "black", tl.srt = 45, diag = F)

v1 <- data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  mutate(usage = if_else(usage <= 4, "No", "Yes")) %>%
  mutate(usage = factor(usage)) %>%
  spread(key="drug", value="usage") %>%
  select(all_of(drug.cols)) %>%
  colpair_map(., cramerV)

v1[, -1][is.na(v1[, -1])] = 1
mat = as.matrix(v1[, -1])
colnames(mat) = colnames(v1[, -1])
rownames(mat) = colnames(v1[, -1])

corrplot(mat, method = "square", type = "lower", order = 'FPC', tl.col = "black", tl.srt = 45, diag = F)


```

> ## density in usgae groups in each drug for each score

```{r  fig.align="center", echo = FALSE,fig.width = 7}

# i = "Alcohol"
# 
# for (i in drug.cols) {
#   plt <- data %>%
#     ggplot(., aes(x=Nscore, group=!!sym(i), fill=!!sym(i))) +
#     geom_density(adjust=1.5, alpha=0.2)
#   
#   print(plt)
# }

# for (i in score.cols) {
#   plt <- data %>%
#     gather(key="drug", value="usage", drug.cols) %>%
#     mutate(usage = as.factor(usage)) %>%
#     ggplot(., aes(x=!!sym(i), group=usage, fill=usage)) +
#     geom_density(adjust=1.5, alpha=0.5) +
#     scale_fill_brewer(palette="Blues") +
#     facet_wrap(~drug)
#   print(plt)
# }


for (i in drug.cols) {
  plt <- data %>%
    gather(key="score", value="val", score.cols) %>%
    mutate(group = ifelse(!!sym(i) >= 4, "active", "inactive"))%>%
    ggplot(., aes(x=val, y=group, fill=group)) +
    geom_joy(scale=1, alpha=0.7, rel_min_height=0.02) +
    labs(title=paste("denstiy plot for", i)) +
    facet_wrap(~score)
  print(plt)
}

```

> ## boxplot distribution in usage groups in each drug for each score

```{r  fig.align="center", echo = FALSE,fig.width = 10, fig.height=7}
for (i in score.cols) {
  plt <- data %>%
    gather(key="drug", value="usage", drug.cols) %>%
    mutate(usage = as.factor(usage)) %>%
    ggplot(., aes(y=!!sym(i), x=usage, fill=usage)) +
    geom_boxplot() +
    scale_fill_brewer(palette="Dark2") +
    facet_wrap(~drug)
  print(plt)
}

```

> ## score changes with usage increase for predifined drug groups

```{r fig.align="center", echo = FALSE,fig.width = 14}
data %>%
  gather(key="drug", value="usage", drug.cols) %>% 
  mutate(drug = case_when(drug %in% hallucinogens ~ "hallucinogens",
                               drug %in% depressants ~ "depressants",
                               drug %in% stimulants ~ "stimulants",
                               T ~ "unknown")) %>%
  group_by(drug, usage) %>%
  summarise_at(vars(score.cols), mean) %>%
  filter(drug != "unknown") %>%
  gather(key="score", value="val", score.cols) %>%
  ggplot(., aes(x=usage, y=val, color=drug)) +
  geom_line(linewidth=1) +
  scale_color_brewer(palette="Set1") +
  facet_wrap(~score)


data %>%
  gather(key="drug", value="usage", drug.cols) %>%
  group_by(drug, usage) %>%
  summarise_at(vars(score.cols), mean) %>%
  filter(drug %in% dng.drugs) %>% 
  group_by(usage) %>%
  summarise_at(vars(score.cols) , mean) %>%
  gather(key="score", value="val", score.cols) %>%
  ggplot(., aes(x=usage, y=val)) +
  geom_line(linewidth=1) +
  scale_color_brewer(palette="Set1") +
  facet_wrap(~score)


  
  
  
```

> score changes with age


```{r}
for (i in score.cols) {
  plt <- data %>%
    ggplot(., aes(y=!!sym(i), x=Age, fill=Age)) +
    geom_boxplot() +
    scale_fill_brewer(palette="Dark2")
  print(plt)
}

```
```{r}
for (i in score.cols) {
  plt <- ggplot(data, aes(y=Education, x=!!sym(i), fill=Education)) +
    geom_boxplot()
  print(plt)
}




```


```{r}

data %>%
  mutate_at(vars(score.cols), ~(. - mean(.))/ sd(.)) %>%
  gather(key="score", value="val", score.cols) %>%
  filter(Crack >= 4) %>%
  ggplot(., aes(x=val)) +
  geom_histogram() +
  facet_wrap(~score)

data %>%
  mutate_at(vars(score.cols), ~ cut(.,
                                    breaks=c(-Inf, -1, 0, 1, Inf),
                                    labels=c("Very Low", "Low", "High", "Very High"),
                                    )
  ) %>%
  gather(key="score", value="level", score.cols) %>%
  mutate(level = factor(level, levels=c("Very Low", "Low", "High", "Very High"))) %>%
  filter(Heroin >= 4) %>%
  ggplot(., aes(y=level)) +
  geom_bar() +
  facet_wrap(~score)



```


> ## MCA

```{r}

dane <- data %>%
  mutate_at(vars(score.cols), ~ cut(.,
                                    breaks=c(-Inf, -1, 0, 1, Inf),
                                    labels=c("Very Low", "Low", "High", "Very High"),
                                    ordered=T)
  ) %>%
  select(-all_of(drug.cols))

  
```


```{r}
res.mca <- MCA(dane)

```


```{r}

res.km = kmeans(res.mca$ind$coord, centers=2, nstart=25, iter.max=50)
fviz_mfa_ind(res.mca, habillage=as.factor(res.km$cluster), palette=c("darkred", "indianred2"), addEllipses=T, repel=T, geom="point")

fviz_nbclust(res.mca$ind$coord, kmeans, method = "silhouette")


```


```{r}
```


```{r}
fviz_mca_biplot(res.mca, 
               repel = TRUE, # Avoid text overlapping (slow if many point)
               ggtheme = theme_minimal(), label = FALSE)

```

```{r}
fviz_mca_var(res.mca, choice = "mca.cor", 
            repel = TRUE, # Avoid text overlapping (slow)
            ggtheme = theme_minimal())

```


```{r}
fviz_mca_var(res.mca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, # Avoid text overlapping
             ggtheme = theme_minimal(),
             select.var = list(cos2 = 0.15))

```


```{r}
# Contributions of rows to dimension 1
fviz_contrib(res.mca, choice = "var", axes = 1, top = 15)

# Contributions of rows to dimension 2
fviz_contrib(res.mca, choice = "var", axes = 2, top = 15)

fviz_contrib(res.mca, choice = "var", axes = 1:2, top = 15)

```


```{r}
fviz_mca_ind(res.mca, col.ind = "cos2", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             ggtheme = theme_minimal(),
             labels=F)

```
```{r}
for (i in colnames(dane)) {
  plt <- dane %>%
    mutate(cluster = res.km$cluster) %>%
    ggplot(aes(x=cluster, fill=!!sym(i))) +
    geom_bar(position="fill")
  print(plt)
}

```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


> ## FAMD

```{r}

cols = c("Country", "Ethnicity")

for (col in cols) {
  features[[col]] = recode_factor(features[[col]], "Other" = paste0(col, "other"))
}

res.famd = FAMD(features)
  

```


```{r}

fviz_famd_var(res.famd, "quanti.var", repel=T, col.var="contrib")

```


```{r}

fviz_famd_var(res.famd, "quali.var", repel=T, col.var="contrib")

```


```{r}

fviz_famd_var(res.famd, "var", repel=T, col.var="coord")

```


```{r}
fviz_screeplot(res.famd)
```


```{r}
as.data.frame(res.famd$eig)

```


```{r}
fviz_mfa_ind(res.famd, geom="point")

res.km = kmeans(res.famd$ind$coord, centers=2, nstart=25, iter.max=50)
fviz_mfa_ind(res.famd, habillage=as.factor(res.km$cluster), palette=c("darkred", "indianred2"), addEllipses=T, repel=T, geom="point")

fviz_nbclust(res.famd$ind$coord, kmeans, method = "silhouette")


```


```{r}
for (i in feature.cols) {
  plt <- features %>%
    mutate(cluster = res.km$cluster) %>%
    ggplot(aes(x=cluster, fill=!!sym(i))) +
    geom_bar(position="fill")
  print(plt)
}


```


```{r}

```


```{r}
```


```{r}
```


```{r}
```

```{r fig.align="center", echo = FALSE,fig.width = 10, fig.height=7}
display.brewer.all()
brewer.pal.info
```
